Flaring star-planet interactions in star-planet systems in Kepler and TESS

### 2023-05-17 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract with filler text
\begin{abstract}
    In many of the star-planet systems discovered so far, most planets orbit far within Mercury's orbit, and many are so close to their host that they plow through its extended magnetosphere. In these systems, planets become in-situ probes of the dynamic magnetic fields in their orbits, able to trigger flares on the stellar surface. Until recently, however, these flaring star-planet interactions suffered from small sample sizes, and poor orbital phase coverage.

@@

    Out of 1811 searched systems, 26 are single star systems that show three or more flares each. For these, we quantified the intensity of excess flaring, and compared it against the expected amplitude of interaction based on the system's parameters. Most systems are neither expected to interact, nor do they show excess flaring. Among those systems that have high expected power of magnetic interaction, we see two branches. An inactive one, where no signs of excess flaring appear regardless of favorable conditions, and an active one where expected power and excess flaring are correlated. While this trend is tentative, it is consistent across all investigated models for magnetic interaction. Finally, we searched for signs of tidal interaction, that is, excess flaring in phase with the tidal bulges ($P_{orb}/2$), and found a similar trend with different expected powers of tidal interaction. Our results support the view that young systems with close in planets are likely to interact both tidally and magnetically, whereas star-planet interactions in old, slowly rotating systems lack the extended magnetosphere, and interact predominantly tidally.
    %magnetic field strength (using $Ro$-$B$ relations), orbital parameters, and planet radius.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits so close to their host star that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Their effects pose challenges, but also opportunities for our understanding of star-planet systems:

On the one hand, a close-in planet that changes the star's rotation and activity makes it more difficult for us to model the system's evolution, as it affects fundamental stellar parameters. Inferring stellar age from rotation may be biased, if the star is spun up or spun down relative to stars without close-in companions~\citep{tejadaarevalo2021further, brown2014discrepancies, maxted2015comparison}. Analogously, altered activity levels may mimic an older or younger host than it actually is~\citep{ilic2022tidal}.

On the other hand, understanding short-term, magnetic star-planet interactions -- the focus of this work -- can provide important insights into the habitability of the entire system~\citep{shkolnik2018signatures}. Even if they orbit far inside the habitable, planets involved in such interactions are in-situ probes of the system's space weather and mass loss by revealing the extent of the Alfv\'en zone~\citep{kavanagh2021planetinduced, chebly2022destination}. Prospectively, planet-induced stellar activity could reveal otherwise difficult to measure planetary properties such as the strength and extent of their magnetic field~\citep{cauley2019magnetic}.

The search for magnetic SPI (Section~\ref{sec:intro:mspi}) in the form of changes in global activity indicators~(Section~\ref{sec:intro:local}), has so far been a series of detections and non-detections, often due to selection bias in the studied systems. Looking for local changes that trace the magnetic footpoint that connects planet to star mitigates this bias, but introduces another -- sampling bias due to poor orbital phase coverage. Flares~(Section~\ref{sec:intro:flares}) are activity markers that can be caused by SPI~(Section~\ref{sec:intro:fspi}), and be monitored more easily with high phase coverage than chromospheric lines or coronal brightness over long periods of time with missions like Kepler and TESS. Thanks to their enormous archives of hundreds of cumulative years of stellar monitoring, we can overcome the limitations of individual system studies~(Section~\ref{sec:intro:local}), and now look for flaring SPI more comprehensively, and systematically.

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic star-planet interactions can only occur if the planet is orbiting within the Alfv\'en radius of the star. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in some way in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby} and \citet{zarka2007plasma, saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. All suffer from a lack of observational constraints for calibration.

\subsection{Searching for global changes in actvity}
\label{sec:intro:global}
Measuring magnetic star planet-interaction is notoriously difficult in individual systems because a non-detection may not mean an absence of the interaction, but just a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive, poppenhaeger2010coronal}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}. Tidal interactions can also induce global increase in activity~\citep{ilic2022tidal}, which adds further ambiguity to the interpretation.

@@

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Currently, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. As will become clear in the following Section, their key advantage for this work is the quasi uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

\subsection{Flaring star-planet interactions}
\label{sec:intro:fspi}
The absolute power of flaring star-planet interactions is difficult to constrain, so we do not know how energetic planet-induced flares can be. The estimates of the maximum dissipated power~\citep{lanza2018closeby} are well within the regime of superflares typically observed by Kepler and TESS. However, on the one hand, that power is an upper limit, so the observed energies might be lower. The planet might for instance be triggering flares prematurely, leading to an overabundance of low energy flares~\citep{loyd2023flares}. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism, so it might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should follow the same process as intrinsic flares.

@@

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable architectures and magnetic field states, the footpoint visibility will be modulated in phase with the planetary orbit. The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
Phase-correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an X-ray active Hot Jupiter host, the observed slight clustering of flaring in orbital phase remains tentative~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed an eruption of X-ray emissionwhen the Jupiter-sized companion was in periastron on its highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner TRAPPIST-1 planets to look for clustered flaring in phase, but found none. \citet{ilin2022searching} and \citet{klein2022one} found hints of excess flaring, and chromospheric Helium emission in the young AU Mic system, respectively. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

In this work, we searched for excess flaring in all available Kepler and TESS short cadence light curves of confirmed star-planet systems~(Section~\ref{sec:data}). We scoured the light curves for flares, and tested their observed phases for departures from a uniform distribution in phase with the innermost planet's orbit~(Section~\ref{sec:methods}). We present the catalog of all flares found in our sample of star-planet systems, and compare the test results with the expected power of interaction for each system~(Section~\ref{sec:results}). We discuss the results for the most interesting star-planet systems individually~(Section~\ref{sec:results:individualstars}), and take a look at the patterns emerging from our analysis~(Section~\ref{sec:discussion}). We summarize our findings in Section~\ref{sec:summary}.

@@

\section{Data}
\label{sec:data}
In this study, we wanted to compile the largest possible sample to search for flaring star-planet interactions. We took the entire Planetary Systems Composite Parameters Table (PSCP, Section~\ref{sec:data:sps})\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}} from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from star-planet interaction, we then required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we estimated the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:orbitalperiod}-\ref{sec:data:lum}). We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table from the literature. The literature parameters for the systems in the final sample of systems, for which we performed the full analysis, is shown in Table~\ref{tab:maintable_lit}.

@@

\subsection{Star-planet systems}
\label{sec:data:sps}
We compiled our sample of star-planet systems from the PSCP, from which we removed controversial planet detections (``pl\_controv\_flag'' set to 1), i.e. detections with existing literature questioning the result. The catalog then contained 2993 transiting and 191 non-transiting systems, from which we picked the innermost known planet in each system.

\subsection{Kepler and TESS photometry}
\label{sec:data:photometry}
The Kepler and TESS missions are unbeaten with respect to long-term optical monitoring of stellar flares. Their excellent coverage of orbital phases makes the light curves ideal for our search for orbital phase dependent flare timing.

Between 2009 and 2013, the Kepler space telescope~\citep{koch2010kepler} nearly continuously observed a patch of the sky in the Cygnus-Lyra region. Each of the 18 observing Quarters contains nearly uninterrupted $\sim 90$ days of observations, totaling over $100\,000$ stars monitored in 2-min cadence in a broad $400-850\,$nm bandpass.
% 100000 from here: https://exoplanetarchive.ipac.caltech.edu/docs/KeplerMission.html

@@

If the eccentricity is known, then the uncertainty on the distance is set to either the error on the semi-major axis, or to half of the difference between periastron and apoastron -- whichever is larger. If eccentricity is not known, then the uncertainty on the distance is either the 25\% error on the semi-major axis, that is, assuming $e=0.25$, or the uncertainty on the semi-major axis -- whichever is larger. We chose $e=0.25$ because it is both a typical value within our sample~(see Table~\ref{tab:maintable_lit}), and in the literature~\citep{eylen2019orbital}.

@@

The rotation of GJ 3082 was measured and are consistent both in TESS and in KELT light curves to below 1\,s precision~\citep{gunther2020stellar}. The rotation of GJ 674 is consistent between light curve periodograms~\citep{kiraga2007agerotationactivity}, and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ 3323 only has rotation periods inferred from activity indicators~\citep{astudillo-defru2017magnetic}, without any periodicity detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation of Proxima Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P/2$, as with GJ 3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ 674.
\end{itemize}

@@

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process, implemented as the \texttt{custom\_detrending} method in \texttt{AltaiPony}, a flare science package for light curve analysis~\citep{ilin2021altaipony}. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ (or $1.5 \sigma$ for very active stars like AU Mic or Proxima Cen) threshold as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked.

@@

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in their light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been covered by the data, also for each light curve separately. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare this distribution to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for potential biases in sensitivity of the test at different phases.

@@

\subsection{Relative velocity}
\label{sec:methods:relvel}
We calculate the relative velocity between the magnetic field of the star and the planet using stellar rotation period $P_{rot}$, orbital period $P_{orb}$, and semi-major axis $a$. We assume that the large scale magnetic field is co-rotating with the stellar surface, and calculate the relative velocity at the planet's orbital distance:

\begin{equation}
    v_{rel} = 2 \pi a \left(\frac{1}{P_{orb}} - \frac{1}{P_{rot}}\right).
\end{equation}

@@

\subsection{Stellar magnetic fields}
\label{sec:methods:bfield}
We derive the average magnetic field strength $B$ from $Ro$ using the empirical relation derived in \cite{reiners2022magnetism},  Table 2, in the unsaturated and saturated regimes, respectively:

\begin{eqnarray}
    B &= 199\,\text{G} \cdot Ro^{-1.26\pm 0.1} \;(\text{if}\; Ro > 0.13) \\
    B &= 2050\,\text{G} \cdot Ro^{-0.11\pm 0.03} \;(\text{if}\; Ro < 0.13) 
\end{eqnarray}

The estimate is consistent with existing Zeeman broadening for, e.g. AU~Mic (3010 G~\cite{reiners2022magnetism}).
\subsection{Power of star-planet interaction}
\label{sec:methods:pspi}
We consider two theories for the mechanism behind flaring star-planet interactions, the stretch-and-break and the Alfv\'en wing mechanisms. For both of them, we include the case of a magnetized and an unmagnetized planet, for a total of four estimates of the expected power of magnetic star-planet interaction.

@@

$B_p$ and $B_*$ are the total planetary and stellar surface field strengths, respectively; $v_{rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_p$ are the stellar and planetary radii, respectively.

@@

Third, for comparison, we estimate the power of star-planet interaction generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, combining Eqs.~8~and~11 in \citet{kavanagh2022radio} to give

@@

@@

We find that no single system shows significant signs of tidal star-planet interaction. Overall, the signal in phase with $P_{orb}/2$ is smaller than with $P_{orb}$. However, in all three scenarios, the sample's deviation from random flare timing increases with expected tidal interaction strength. We can compare the magnetic interaction measurements to the tidal ones in some of the conspicuous systems:

>>>>

% Abstract with filler text
\begin{abstract}
    In many of the star-planet systems discovered so far, the innermost planets orbit extremely close to their host, that is, within its extended magnetosphere. In these systems, the planets become in-situ probes of the stellar magnetic fields. As they move, they may trigger flares on the stellar surface. Until recently, searches for such flaring star-planet interactions suffered from small sample sizes, and poor orbital phase coverage.

@@

    In the 1811 searched systems, we found 26 single stars with three or more flares each. We quantified the significance of their excess flaring, and compared it against the theoretically expected amplitude of magnetic interaction. Most systems are neither expected to, nor do they show excess flaring. Among the systems that we can expect to interact, we see two branches. An inactive one, where no signs of excess flaring appear, and an active one, where expected power and excess flaring are correlated. This trend is tentative, but consistent for all scaling laws for magnetic interaction.

    Finally, we searched the same data for signs of locally expressed tidal interaction, that is, excess flaring in phase with the tidal bulges ($P_{\rm orb}/2$). We found a similar trend with different scaling laws of tidal interaction, albeit at even lower significance. Our system-by-system comparison suggests that young systems may interact both tidally and magnetically, whereas old, magnetically inactive systems might mostly interact tidally.
    %magnetic field strength (using $Ro$-$B$ relations), orbital parameters, and planet radius.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits so close to their host star that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Specifically, the effects on the star pose challenges, but also opportunities for our understanding of star-planet systems:

On the one hand, a close-in planet that changes the star's rotation and activity makes it more difficult for us to model the system's evolution, as it affects fundamental stellar parameters. Ages inferred from rotation will be biased, if the star is spun up or spun down relative to stars without close-in companions~\citep{tejadaarevalo2021further, brown2014discrepancies, maxted2015comparison}. Analogously, altered activity levels may mimic an older or younger host than it actually is~\citep{ilic2022tidal}. On the other hand, measuring in what ways and how much close-in planets alter their hosts' behavior can provide important insights into the habitability of the entire system~\citep{shkolnik2018signatures}. Even if they orbit far inside the habitable zone, such planets probe the system's space weather and mass loss by revealing the extent of the Alfv\'en zone~\citep{kavanagh2021planetinduced, chebly2022destination}. % Prospectively, planet-induced stellar activity could quantify otherwise difficult to measure planetary properties such as the strength and extent of their magnetic field~\citep{cauley2019magnetic}.

The search for magnetic SPI (Section~\ref{sec:intro:mspi}) in the form of changes in global activity indicators~(Section~\ref{sec:intro:global}), has so far been a series of detections and non-detections, often due to selection bias in the studied systems. Looking for local changes that trace the magnetic footpoint that connects planet to star mitigates this bias, but introduces another -- sampling bias due to poor orbital phase coverage. Flares~(Section~\ref{sec:intro:flares}) are activity markers that can be caused by SPI~(Section~\ref{sec:intro:fspi}). They can be monitored with high phase coverage over long periods of time with missions like Kepler and TESS. Thanks to their large archives of hundreds of cumulative years of stellar monitoring, we can overcome the limitations of individual system studies~(Section~\ref{sec:intro:local}), and now look for flaring SPI more comprehensively, and systematically.

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic star-planet interactions can only occur if the planet is orbiting within the Alfv\'en radius of the star. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints for calibration, which we aim to provide in this work.

\subsection{Searching for global changes in activity}
\label{sec:intro:global}
Measuring magnetic star planet-interaction is notoriously difficult in individual systems because a non-detection may not mean an absence of the interaction, but merely a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive, poppenhaeger2010coronal}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}. Tidal interactions can also induce global increase in activity~\citep{ilic2022tidal}, which adds further ambiguity to the interpretation.

@@

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Currently, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. Their key advantage for this work is the nearly uninterrupted monitoring of star-planet systems that covers all phases of a close-in planet's orbit.

\subsection{Flaring star-planet interactions}
\label{sec:intro:fspi}
The absolute power of flaring star-planet interactions is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates of the maximum dissipated power~\citep{lanza2018closeby} are well within the regime of superflares typically observed by Kepler and TESS. However, on the one hand, that power is an upper limit, so the observed energies might be lower. The planet might, for instance, be triggering flares prematurely, leading to an overabundance of low energy flares~\citep{loyd2023flares}. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should follow the same process as intrinsic flares.

@@

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable viewing geometries and magnetic field states, the footpoint visibility will be modulated in phase with the planetary orbit. The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
In contrast with global SPI, local SPI effects are easier to tell apart from intrinsic stellar behavior because they are tied to the planet's orbital period. 
Phase-correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an X-ray active Hot Jupiter host, the observed slight clustering of flaring in orbital phase remains tentative~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed an eruption of X-ray emission when the Jupiter-sized companion was in periastron on its highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner TRAPPIST-1 planets to look for clustered flaring in phase, but found none. \citet{ilin2022searching} and \citet{klein2022one} found hints of excess flaring, and chromospheric Helium emission in the young AU Mic system, respectively. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

\subsection{Overview}
In this work, we searched for excess flaring in all available Kepler and TESS short cadence light curves of confirmed star-planet systems~(Section~\ref{sec:data}). We scoured the light curves for flares, and tested their observed phases for departures from a uniform distribution in phase with the innermost planet's orbit~(Section~\ref{sec:methods}). We present the catalog of all flares found in our sample of star-planet systems, and compare the test results with the expected power of interaction for each system~(Section~\ref{sec:results}). We discuss the results for the most interesting star-planet systems individually~(Section~\ref{sec:results:individualstars}), and take a look at the patterns emerging from our analysis~(Section~\ref{sec:discussion}), particularly at the possible role of tidal flaring SPI in our observations~(Section~\ref{sec:discussion:tidal}).  We summarize our findings in Section~\ref{sec:summary}.

@@

\section{Data}
\label{sec:data}
In this study, we compiled the largest possible sample to search for flaring star-planet interactions. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from star-planet interaction, we then sought out the orbital periods of each system from the literature~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we estimated the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we also mostly found in the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date results. The parameters for the systems in the final sample of systems are shown in Table~\ref{tab:maintable_lit}.

@@

\subsection{Star-planet systems}
\label{sec:data:sps}
We compiled our sample of star-planet systems from the PSCP, from which we removed controversial planet detections (``pl\_controv\_flag'' set to 1), i.e., detections with existing literature questioning the result. The resulting catalog contained 2993 transiting and 191 non-transiting systems, from which we picked the innermost known planet in each system.

\subsection{Kepler and TESS photometry}
\label{sec:data:photometry}
The Kepler and TESS missions are unbeaten with respect to long-term optical monitoring of stellar flares. Their excellent coverage of orbital phases makes the light curves ideal for our search for orbital phase dependent flare timing. Between 2009 and 2013, the Kepler space telescope~\citep{koch2010kepler} nearly continuously observed a patch of the sky in the Cygnus-Lyra region. Each of the 18 observing Quarters contains nearly uninterrupted $\sim 90$ days of observations, totaling over $100\,000$ stars monitored in 2-min cadence in a broad $400-850\,$nm bandpass.
% 100000 from here: https://exoplanetarchive.ipac.caltech.edu/docs/KeplerMission.html

@@

If the eccentricity is known, then the uncertainty on the distance is set to either the error on the semi-major axis, or to half of the difference between periastron and apoastron -- whichever is larger. If eccentricity is not known, then the uncertainty on the distance is set to either the 25\% error on the semi-major axis, that is, assuming $e=0.25$, or the uncertainty on the semi-major axis -- whichever is larger. We chose $e=0.25$ because it is both a typical value within our sample~(see Table~\ref{tab:maintable_lit}), and in the literature~\citep{eylen2019orbital}.

@@

Among the stars without given uncertainties, the rotation of GJ 3082 was measured and is consistent both in TESS and in KELT light curves to below 1\,s precision~\citep{gunther2020stellar}. The rotation of GJ 674 is consistent between light curve periodograms~\citep{kiraga2007agerotationactivity}, and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ 3323 only has rotation periods inferred from activity indicators~\citep{Astudillo-defru2017magnetic}, without any periodicity detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation of Proxima Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P/2$, as with GJ 3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ 674.
\end{itemize}

@@

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process, implemented as the \texttt{custom\_detrending} method in \texttt{AltaiPony}, a flare science package for light curve analysis~\citep{ilin2021altaipony}. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ (or $1.5 \sigma$ for very active stars like AU Mic or Proxima Cen) threshold as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked. Transits are usually too shallow to affect either light curve de-trending or flare finding. The light curve portions around deep transits were inspected manually.

@@

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been covered by the data, also for each light curve separately. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare this distribution to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for potential biases in sensitivity of the test at different phases.

@@

\subsection{Relative velocity}
\label{sec:methods:relvel}
We calculate the relative velocity between the magnetic field of the star and the planet using stellar rotation period $P_{rot}$, orbital period $P_{\rm orb}$, and semi-major axis $a$. We assume that the large scale magnetic field is co-rotating with the stellar surface, and calculate the relative velocity at the planet's orbital distance:

\begin{equation}
    v_{rel} = 2 \pi a \left(\frac{1}{P_{\rm orb}} - \frac{1}{P_{rot}}\right).
\end{equation}

@@

\subsection{Stellar magnetic fields}
\label{sec:methods:bfield}
We derive the average magnetic field strength $B$ from $R$o using the empirical relation derived in \citet{reiners2022magnetism},  Table 2, in the unsaturated and saturated regimes, respectively:

\begin{eqnarray}
    B &= 199\,\text{G} \cdot R\rm o^{-1.26\pm 0.1} \;(\text{if}\; Ro > 0.13) \\
    B &= 2050\,\text{G} \cdot R\rm o^{-0.11\pm 0.03} \;(\text{if}\; Ro < 0.13) 
\end{eqnarray}

The convective turnover time $\tau$ in  $R \rm o=P_{rot}/\tau$ is derived using bolometric luminosity~(Section~\ref{sec:data:lum}), following~\citet{reiners2014generalized, reiners2022magnetism}:

\begin{equation}
    \tau = 12.3\, \rm d \cdot (L_{\rm bol}/L_\odot)^{-1/2}
\end{equation}

\subsection{Power of star-planet interaction}
\label{sec:methods:pspi}
We consider two theories for the mechanism behind flaring star-planet interactions, the stretch-and-break and the Alfv\'en wing mechanisms. For both of them, we include the case of a magnetized and an unmagnetized planet, for a total of four estimates of the expected power of magnetic star-planet interaction.

@@

$B_p$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_p$ are the stellar and planetary radii, respectively. Since the planetary fields are unknown, we ignore it in the calculations. Since we also do not know the stellar field strength at the pole for the vast majority of the stars, we adopt the surface field strength~(Section~\ref{sec:methods:bfield}) as a proxy.

@@

Third, for comparison, we estimate the power of star-planet interaction generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, combining Eqs.~8~and~11 in \citet{kavanagh2022radio} to arrive at

@@

Lacking $B_p$ and various additional factors that appear in $P_{sb},P_{sb0},P_{aw}$, and  $P_{aw0}$, such as the angle between the wind and the

@@

We find that no single system shows significant signs of tidal star-planet interaction. Overall, the signal in phase with $P_{\rm orb}/2$ is smaller than with $P_{\rm orb}$. However, in all three scenarios, the sample's deviation from random flare timing increases with expected tidal interaction strength. We can compare the magnetic interaction measurements to the tidal ones in some of the conspicuous systems:
